Bipartite Fidelity and Loschmidt Echo of the Bosonic Conformal Interface

### 2017-12-13 F_correction.tex
[prb] version accepted by prb.

% outer boundary correction
It is discussed in Cardy and Peschel's work\cite{cardy_finite-size_1988} that boundary will contribute a logarithmic term in the free energy,
\begin{equation}
F = - \frac{c}{6} \left( \int_M  K(x) d^2x + \int_{\partial M} k_g ds \right)  \ln L , 
\end{equation}
where $M$ is a 2d smooth manifold, $K(x)$ is the Gaussian curvature, $k_g$ is the geodesic curvature of the boundary of the manifold and $L$ is the system's characteristic length.

@@

The same thing happened in Fig.~\ref{fig:H-tau_fold} with a slightly different mechanism. In the slit diagram (left panel in Fig.~\ref{fig:H-tau_fold}), the regulators all have radii that are at the order of the short distance cut-off. They will have negligible contributions to the free energy. However, in the new $\xi$ plane, we implicitly switch to a {\it new short distance regulator} such that only the blue semi-circle around $0$ contribute negligibly. The outer surface radius, despite being the image of a small semi-circle on $z$ plane, will contribute a $-\frac{c}{6} \ln \frac{r_2}{a}$ term on the $\xi$ plane that should not be there.

>>>>

% outer boundary correction
It is discussed in Cardy and Peschel's work\cite{cardy_finite-size_1988} that the boundary will contribute logarithmic term in the free energy,
\begin{equation}
F = - \frac{c}{6} \left( \int_M  K(x) d^2x + \int_{\partial M} k_g ds \right)  \ln L , 
\end{equation}
where $M$ is a 2d smooth manifold, $K(x)$ is the Gaussian curvature, $k_g$ is the geodesic curvature of the boundary of the manifold and $L$ is the system's characteristic length.

@@

The same thing happened in Fig.~\ref{fig:H-tau_fold} with a slightly different mechanism. In the slit diagram (left panel in Fig.~\ref{fig:H-tau_fold}), the regulators all have radii that are at the order of the short distance cut-off. They will have negligible contributions to the free energy. However, in the new $\xi$ plane, we implicitly switch to a {\it new short distance regulator} such that only the blue semi-circle around $0$ contributes negligibly. The outer surface radius, despite being the image of a small semi-circle on the $z$ plane, will contribute a $-\frac{c}{6} \ln \frac{r_2}{a}$ term on the $\xi$ plane that should not be there.

### 2017-12-13 abstract.tex
[prb] version accepted by prb.

We study the quantum quench problem for a class of bosonic conformal interfaces by computing the Loschmidt echo and the bipartite fidelity. The quench can be viewed as a sudden change of boundary conditions parameterized by $\theta$ when connecting two one-dimensional critical systems. They are classified by $S(\theta)$ matrices associated with the current scattering processes on the interface. The resulting Loschmidt echo of the quench has long time algebraic decay $t^{-\alpha}$, whose exponent also appears in the finite size bipartite fidelity as $L^{-\frac{\alpha}{2}}$. We perform analytic and numerical calculations of the exponent $\alpha$, and find that it has a quadratic dependence on the change of $\theta$ if the prior and post quench boundary conditions are in the same type of $S$, while remains $\frac{1}{4}$ otherwise. Possible physical realizations of these interfaces include for instance connecting different quantum wires (Luttinger liquids), quench of the topological phase edge states \etc and the exponent can be detected in a X-ray edge singularity type experiment.

>>>>

We study the quantum quench problem for a class of bosonic conformal interfaces by computing the Loschmidt echo and the bipartite fidelity. The quench can be viewed as a sudden change of boundary conditions parameterized by $\theta$ when connecting two one-dimensional critical systems. They are classified by $S(\theta)$ matrices associated with the current scattering processes on the interface. The resulting Loschmidt echo of the quench has long time algebraic decay $t^{-\alpha}$, whose exponent also appears in the finite size bipartite fidelity as $L^{-\frac{\alpha}{2}}$. We perform analytic and numerical calculations of the exponent $\alpha$, and find that it has a quadratic dependence on the change of $\theta$ if the prior and post quench boundary conditions are of the same type of $S$, while remaining $\frac{1}{4}$ otherwise. Possible physical realizations of these interfaces include, for instance, connecting different quantum wires (Luttinger liquids), quench of the topological phase edge states, \etc and the exponent can be detected in a X-ray edge singularity type experiment.

### 2017-12-13 analy_eval.tex
[prb] version accepted by prb.

The Loschmidt echo can be evaluated in the same way. Again, we introduce two semi-circles (blue in Fig.~\ref{fig:H-tau_fold}) as regulators and then perform the conformal transformation shown in Fig.~\ref{fig:H-tau_fold}. From $z$ plane to the $\xi$ plane, we use $\xi = \frac{z}{\tau - z}$ to map the two slits to half of an annulus, which is the same as the fidelity case. With one more conformal mapping $w = \ln \xi$, the diagram again becomes the cylinder partition function between the two boundary states.

@@

One subtlety of the above description is that the two semi-circles in the center diagrams of Fig.~\ref{fig:fidel-map} and Fig.~\ref{fig:H-tau_fold} are not precisely concentric. This can be resolved by the following observation. There exists a conformal map $\zeta(\xi)$ that maps the non-concentric circles to the two standard concentric circles of radii $1$ and $R$ ($R>1$) on the $\zeta$ plane\cite{brown_complex_2009}. Then the logarithmic map $w = \ln \zeta$ produces a cylinder of width $\ln R$. In our case, since the height of the cylinder is always $\pi$, the width of the cylinder is a conformal invariant that only depends on the cross ratio of the half annulus. The four intersection points of two standard concentric circles on $\zeta$ plane are $(\pm 1,0)$ and $(\pm R,0)$, whose cross ratio is
\begin{equation}
\eta = \frac{(1 + R)^2}{(1 - R)^2}. 
\end{equation}
Hence the width of the cylinder is $\ln \frac{\sqrt{ \eta } - 1}{\sqrt{ \eta} + 1}$. Since conformal transformation preserves the cross ratio, the result is the same if we use the cross ratio of the slightly non-concentric diagrams of Fig.~\ref{fig:fidel-map} and Fig.~\ref{fig:H-tau_fold}. The calculation in Eq.~\eqref{eq:fidel_cyd_width} and Eq.~\eqref{eq:echo_cyd_width} equivalently use the leading order approximation to $\eta$ in the respective geometries and thus get the leading order term in the width of the cylinders. The slight deviation to the precise concentric geometry will only bring in $\frac{\epsilon}{L}, \frac{\epsilon}{\tau}$ corrections to $\eta$ and the width parameter, which will not affect the fidelity and echo exponents.

@@

As analyzed in Sec.~\ref{sec:analytic_numerics}, $S_2( \theta)$ interpolates between DD and NN, $S_1( \theta )$ interpolates between DN and ND. In the region accessible to the numerical calculation in the lattice model, we choose the process ${\rm DD} \rightarrow  \lambda$ to verify
\begin{equation}
\label{eq:result_DDDD}
\mathcal{F} = 
\left\lbrace
\begin{aligned}
\frac{1}{8}\ln L  &\quad\text{fidelity}  \\
\frac{1}{4}\ln t   &\quad \text{echo} .  \\
\end{aligned} \right.  
\end{equation}
The same results have already been obtained for $\lambda = {\rm P}$\cite{stephan_logarithmic_2013,stephan_local_2011,vasseur_universal_2014,vasseur_crossover_2013,kennes_universal_2014}. Another process $\rm{DN} \rightarrow \lambda$ {\iffalse \color{red} in Eq.~\eqref{eq:DNDN}\fi} is used to verify
\begin{equation}
\label{eq:result_DNDN}
\mathcal{F} = 
\left\lbrace
\begin{aligned}
 (x - x^2 )\ln L   &  \quad {\rm fidelity} \\
 2(x - x^2 ) \ln t  & \quad \text{ Loschmidt echo} , \\
\end{aligned} \right. 
\end{equation}
where $\lambda = \tan \theta$ and $x = \frac{\theta}{\pi}$.

>>>>

The Loschmidt echo can be evaluated in the same way. Again, we introduce two semi-circles (blue in Fig.~\ref{fig:H-tau_fold}) as regulators and then perform the conformal transformation shown in Fig.~\ref{fig:H-tau_fold}. From the $z$ plane to the $\xi$ plane, we use $\xi = \frac{z}{\tau - z}$ to map the two slits to half of an annulus, which is the same as the fidelity case. With one more conformal mapping $w = \ln \xi$, the diagram again becomes the cylinder partition function between the two boundary states.

@@

One subtlety of the above description is that the two semi-circles in the center diagrams of Fig.~\ref{fig:fidel-map} and Fig.~\ref{fig:H-tau_fold} are not precisely concentric. This can be resolved by the following observation. There exists a conformal map $\zeta(\xi)$ that maps the non-concentric circles to the two standard concentric circles of radii $1$ and $R$ ($R>1$) on the $\zeta$ plane\cite{brown_complex_2009}. Then the logarithmic map $w = \ln \zeta$ produces a cylinder of width $\ln R$. In our case, since the height of the cylinder is always $\pi$, the width of the cylinder is a conformal invariant that only depends on the cross ratio of the half annulus. The four intersection points of two standard concentric circles on the $\zeta$ plane are $(\pm 1,0)$ and $(\pm R,0)$, whose cross ratio is
\begin{equation}
\eta = \frac{(1 + R)^2}{(1 - R)^2}. 
\end{equation}
Hence the width of the cylinder is $\ln \frac{\sqrt{ \eta } - 1}{\sqrt{ \eta} + 1}$. Since conformal transformation preserves the cross ratio, the result is the same if we use the cross ratio of the slightly non-concentric diagrams of Fig.~\ref{fig:fidel-map} and Fig.~\ref{fig:H-tau_fold}. The calculation in Eq.~\eqref{eq:fidel_cyd_width} and Eq.~\eqref{eq:echo_cyd_width} equivalently use the leading order approximation to $\eta$ in the respective geometries and thus get the leading order term in the width of the cylinders. The slight deviation to the precise concentric geometry will only bring in $\frac{\epsilon}{L}, \frac{\epsilon}{\tau}$ corrections to $\eta$ and the width parameter, which will not affect the fidelity and echo exponents.

@@

As analyzed in Sec.~\ref{sec:analytic_numerics}, $S_2( \theta)$ interpolates between DD and NN, and $S_1( \theta )$ interpolates between DN and ND. In the region accessible to the numerical calculation in the lattice model, we choose the process ${\rm DD} \rightarrow  \lambda$ to verify
\begin{equation}
\label{eq:result_DDDD}
\mathcal{F} = 
\left\lbrace
\begin{aligned}
\frac{1}{8}\ln L  &\quad\text{fidelity}  \\
\frac{1}{4}\ln t   &\quad \text{echo} .  \\
\end{aligned} \right.  
\end{equation}
The same results have already been obtained for $\lambda = {\rm P}$\cite{stephan_logarithmic_2013,stephan_local_2011,vasseur_universal_2014,vasseur_crossover_2013,kennes_universal_2014}. Another process $\rm{DN} \rightarrow \lambda$ {\iffalse \color{red} in Eq.~\eqref{eq:DNDN}\fi} is used to verify
\begin{equation}
\label{eq:result_DNDN}
\mathcal{F} = 
\left\lbrace
\begin{aligned}
 (x - x^2 )\ln L   &  \quad {\rm fidelity} \\
 2(x - x^2 ) \ln t  & \quad \text{ Loschmidt echo} , \\
\end{aligned} \right. 
\end{equation}
where $\lambda = \tan \theta$ and $x = \frac{\theta}{\pi}$.

### 2017-12-13 bd_notation.tex
[prb] version accepted by prb.

We use chemical reaction style to represent the change of boundary conditions. Taking the example of the echo diagrams in Fig.~\ref{fig:echo}, there are three boundary conditions $a,b,c$ in the folding picture, which represents the status of the two ends of the chain before and after the quench. The choice of a uniform $c$ boundary condition on the far end of the chain is to isolate the effect coming from the bcc on the $ab$ interface. The process $a + c \rightarrow b + c$ represents the change of boundary condition from the combination $a$/$c$ prior the quench to $b$/$c$ after the quench. Since each letter can take a general conformal interface defined by the $S$ matrix in Eq.~\eqref{eq:S1_S2}, we denote it as
\begin{equation}
\label{eq:Full_notation_rand()}
S_a( \theta_a ) + S_c( \theta_c) \rightarrow S_b( \theta_b )  + S_c( \theta_c ) .
\end{equation}
In most cases of the following, we will consider taking $a = c$ to remove the bcc operator from $a$ to $c$ at infinity. And we will use the shorthand notation
\begin{equation}
S_a( \theta_a ) \rightarrow S_b( \theta_b )
\end{equation}
to remind ourselves that we are isolating the boundary condition change only on the joint of the two chains.

>>>>

We use chemical reaction style to represent the change of boundary conditions. Taking the example of the echo diagrams in Fig.~\ref{fig:echo}, there are three boundary conditions $a,b,c$ in the folding picture, which represents the status of the two ends of the chain before and after the quench. The choice of a uniform $c$ boundary condition on the far end of the chain is to isolate the effect coming from the bcc on the $ab$ interface. The process $a + c \rightarrow b + c$ represents the change of boundary condition from the combination $a$/$c$ prior to the quench to $b$/$c$ after the quench. Since each letter can take a general conformal interface defined by the $S$ matrix in Eq.~\eqref{eq:S1_S2}, we denote it as
\begin{equation}
\label{eq:Full_notation_rand()}
S_a( \theta_a ) + S_c( \theta_c) \rightarrow S_b( \theta_b )  + S_c( \theta_c ) .
\end{equation}
In most cases of the following, we will consider taking $a = c$ to remove the bcc operator from $a$ to $c$ at infinity. And we will use the shorthand notation
\begin{equation}
S_a( \theta_a ) \rightarrow S_b( \theta_b )
\end{equation}
to remind ourselves that we are isolating the boundary condition change only on the joint of the two chains.

### 2017-12-13 comp_fid_echo.tex
[referee] version resubmitted to prb.
[prb] version accepted by prb.

During the course of derivation in this and other appendices, we will repeatedly use the combinatorial identity called McMahon Master theorem
\begin{equation}
\label{eq:bosonic_McMahon}
\langle0|\exp\left\{\frac{1}{2}b_iX_{ij}b_j\right\}\exp\left\{\frac{1}{2}b^\dagger_iY_{ij}b^\dagger_j\right\}|0\rangle=\text{det}^{-\frac{1}{2}}(1-XY),
\end{equation}
for symmetric matrix $X$ and $Y$ and set of independent bosonic creation operators $b_i^{\dagger}$. One can prove it for (simultaneously) diagonalizable matrices and then claim its legitimacy for its combinatorial nature.

\subsection{Boson Bogoliubov transformation}
\label{app_sub:boson_BdG}
We consider the following quadratic bosonic Hamiltonian
\begin{equation}
\begin{aligned}
\label{eq:quadratic_boson_H}
\hat{H} &= \frac{1}{2} (\vec{b}^{\dagger}, -\vec{b}) M 
\begin{pmatrix}
\vec{b}\\
\vec{b}^{\dagger} 
\end{pmatrix},  \qquad 
M = 
\begin{bmatrix}
A & -B^* \\
B & -A^* \\
\end{bmatrix},
\end{aligned}
\end{equation}
where ${\bf b}\equiv(b_{1},...,b_{n})^T$ is a vector of bosonic annihilation operators. The matrix $M$ consists of a $n\times n$ Hermitian block $A$ which plays the role of single particle Hamiltonian in the fermionic case and symmetric block $B$ of the pairing interaction.

@@

To proceed, we introduce the real basis 
\begin{equation}\begin{aligned}
\label{eq:real_basis}
\begin{pmatrix}
{\bf b}\\
{\bf b}^\dagger
\end{pmatrix}
=C\begin{pmatrix}
\phi\\
\pi
\end{pmatrix}
=\frac{1}{\sqrt{2}}\begin{bmatrix}
1 & i \\
1 & -i
\end{bmatrix}\begin{pmatrix}
\phi\\
\pi
\end{pmatrix},
\end{aligned}\end{equation}
in which the Hamiltonian reads
\begin{equation}\begin{aligned}
\hat{H}
&=\frac{1}{2}
\begin{pmatrix}
\phi & \pi
\end{pmatrix}
\begin{bmatrix}
\text{Re}(A-B^*) & -\text{Im}(A)+\text{Im}(B)\\
\text{Im}(A)+\text{Im}(B)& \text{Re}(A+B) \\
\end{bmatrix}
\begin{pmatrix}
\phi\\
\pi
\end{pmatrix} \\
&=\frac{1}{2}
\begin{pmatrix}
\phi & \pi
\end{pmatrix}
\mathcal{M}
\begin{pmatrix}
\phi\\
\pi
\end{pmatrix}.
\end{aligned}\end{equation}
It is not hard to check that $\mathcal{M}$ is real and symmetric.

@@

We will not elaborate on Williamson's theorem and its proof (see proofs in Ref.~\onlinecite{xiao_theory_2009,pirandola_correlation_2009,gosson_symplectic_2006} and also a recent application in entanglement entropy context\cite{coser_contour_2017}). Instead we will show in App.~\ref{app_sub:harmonic_chain} that for the problem of harmonic chain we are interested in, the diagonalization can be easily done without using the general recipe in the Williamson theorem.

@@

Suppose we have obtained the required matrix $S$, the ground state will be the vacuum of the annihilation operators defined in Eq.~\eqref{eq:def_a} and in $b$ basis it satisfies
\begin{equation}\begin{aligned}
\label{eq:a_vacuum_condition}
(b_iu_{ij}+b_i^\dagger v_{ij})|0\rangle_{{\bf a}}=0.
\end{aligned}\end{equation}
If the matrix $u$ is invertible, then we can introduce a matrix $T=vu^{-1}$ to rewrite Eq.~\eqref{eq:a_vacuum_condition} as
\begin{equation}\begin{aligned}
(b_i+b_j^\dagger T_{ji})|0\rangle_{{\bf a}}=0. 
\end{aligned}\end{equation}
The constraint Eq.~\eqref{eq:constraint_2} on the blocks of $u$ and $v$ (followed by the symplectic constraint of $S$) implies that $T$ is a symmetric matrix. With the observation of 
\begin{equation}\begin{aligned}
\exp\left\{-\frac{1}{2}b_j^\dagger T_{jk}b_k^\dagger\right\}b_i\exp\left\{\frac{1}{2}b_j^\dagger T_{jk}b_k^\dagger\right\}=b_i+T_{ij}b^\dagger_j,
\end{aligned}\end{equation}
we solve the groundstate
\begin{equation}\begin{aligned}
\label{eq:boson_BCS_gnd}
|0\rangle_{{\bf a}}&=\text{det}^{\frac{1}{4}}(1-T^\dagger T)\exp\left\{-\frac{1}{2}b_j^\dagger T_{jk}b_k^\dagger\right\}|0\rangle_{\bf b},\\
\end{aligned}\end{equation}
where the normalization is given by the McMahon Master theorem Eq.~\eqref{eq:bosonic_McMahon}. Apply constraint in Eq.~\eqref{eq:constraint_1}, it simplifies to the top left corner of the symplectic matrix
\begin{equation}
\text{det}^{\frac{1}{4}}(1-T^\dagger T) =|\text{det}(u)|^{-\frac{1}{2}}.
\end{equation}

@@

We start with a quadratic bosonic Hamiltonian $\hat{H}_0$ in ${\bf b}$ basis, as in Eq.~\eqref{eq:quadratic_boson_H}. From the discussion in App.~\ref{app_sub:boson_BdG}, we are able to diagonalize it in ${\bf a}$ basis for positive definite $M$. At $t=0$, the Hamiltonian becomes $\hat{H}_1$, which is still written in ${\bf b}$ basis, but is diagonalized in a new basis ${\bf c}$. The corresponding Bogoliubov transformations read
\begin{equation}\begin{aligned}
\label{eq:two_BdG}
({\bf b} , {\bf b}^\dagger)S_0=({\bf a} , {\bf a}^\dagger),\quad
({\bf b} , {\bf b}^\dagger)S_1=({\bf c} , {\bf c}^\dagger),
\end{aligned}\end{equation}
and so
\begin{equation}\begin{aligned}
\label{eq:S0invS}
({\bf a} , {\bf a}^\dagger)=({\bf c} , {\bf c}^\dagger)\left(S_0^{-1}S_1\right)^{-1}.
\end{aligned}\end{equation}
One realizes that Eq.~\eqref{eq:S0invS} is another Bogoliubov transformation and so the corresponding matrix has the block structure
\begin{equation}
S_1^{-1}S_0=\begin{bmatrix}
u_1 & v_1^*\\
v_1 & u_1^*
\end{bmatrix}.
\end{equation}
Thus $|0\rangle_{\bf c}$ is related to the $|0\rangle_{{\bf a}}$ in the same way as in Eq.~\eqref{eq:boson_BCS_gnd}. Their overlap is therefore given by the normalization factor
\begin{equation}\begin{aligned}
|{}_{\bf a}\langle0|0\rangle_{\bf c}|^2=\frac{\Big|{}_{\bf a}\langle 0 | \exp( -\frac{1}{2} a_j^{\dagger} T^{jk} a_k^{\dagger} )|0   \rangle_{\bf a} \Big|^2}{|\text{det}(u_1)|} = \frac{1}{|\text{det}(u_1)|}.
\end{aligned}\end{equation}

@@

In this subsection, we explicitly construct the matrix $\mathcal{S}$ for the case of harmonic chain introduced in Sec.~\ref{sec_sub:free_boson_lattice}. In the basis defined in Eq.~\eqref{eq:real_basis}, the Hamiltonian for 1D harmonic chain reads
\begin{equation}\begin{aligned}
\hat{H}
=\frac{1}{2}
\begin{pmatrix}
\phi & \pi
\end{pmatrix}
\mathcal{M}
\begin{pmatrix}
\phi\\
\pi
\end{pmatrix}
=\frac{1}{2}
\begin{pmatrix}
\phi & \pi
\end{pmatrix}
\begin{bmatrix}
\mathcal{V} \\
& {\mathbb{ I}}
\end{bmatrix}
\begin{pmatrix}
\phi\\
\pi
\end{pmatrix},
\end{aligned}\end{equation}
where $\mathcal{V}$ is real symmetric matrix that can be diagonalized as $\mathcal{V}=\mathcal{O}D^2\mathcal{O}^T$. The matrix $\mathcal{V}$ depends on the boundary condition, but positive definiteness is the only requirement here.

The matrix $\mathcal{S}$ that diagonalize $\mathcal{M}$,
\begin{equation}
\begin{aligned}
\mathcal{M}=\mathcal{S}
\begin{bmatrix}
D \\ 
& D
\end{bmatrix}
\mathcal{S}^T
\end{aligned}
\end{equation}
is given by the following real symplectic matrix
\begin{equation}\begin{aligned}
\mathcal{S}\equiv
\begin{bmatrix}
\mathcal{O}D^{1/2} \\
& \mathcal{O}D^{-1/2}
\end{bmatrix}.
\end{aligned}\end{equation}

>>>>

During the course of derivation in this and other appendices, we will repeatedly use the combinatorial identity called the McMahon master theorem
\begin{equation}
\label{eq:bosonic_McMahon}
\langle0|\exp\left\{\frac{1}{2}b_iX_{ij}b_j\right\}\exp\left\{\frac{1}{2}b^\dagger_iY_{ij}b^\dagger_j\right\}|0\rangle=\text{det}^{-\frac{1}{2}}(1-XY),
\end{equation}
for symmetric matrix $X$ and $Y$ and set of independent bosonic creation operators $b_i^{\dagger}$. One can prove it for (simultaneously) diagonalizable matrices and then claim its legitimacy for its combinatorial nature.

\subsection{Boson Bogoliubov transformation}
\label{app_sub:boson_BdG}
We consider the following quadratic bosonic Hamiltonian
\begin{equation}
\begin{aligned}
\label{eq:quadratic_boson_H}
\hat{H} &= \frac{1}{2} (\vec{b}^{\dagger}, -\vec{b}) M 
\begin{pmatrix}
\vec{b}\\
\vec{b}^{\dagger} 
\end{pmatrix},  \qquad 
M = 
\begin{bmatrix}
A & -B^* \\
B & -A^* \\
\end{bmatrix},
\end{aligned}
\end{equation}
where ${\bf b}\equiv(b_{1},...,b_{n})^T$ is a vector of bosonic annihilation operators. The matrix $M$ consists of a $n\times n$ Hermitian block $A$ which plays the role of a single particle Hamiltonian in the fermionic case and symmetric block $B$ of the pairing interaction.

@@

To proceed, we introduce the real basis 
\begin{equation}\begin{aligned}
\label{eq:real_basis}
\begin{pmatrix}
{\bf b}\\
{\bf b}^\dagger
\end{pmatrix}
=C\begin{pmatrix}
\phi\\
\pi
\end{pmatrix}
=\frac{1}{\sqrt{2}}\begin{bmatrix}
1 & i \\
1 & -i
\end{bmatrix}\begin{pmatrix}
\phi\\
\pi
\end{pmatrix},
\end{aligned}\end{equation}
in which the Hamiltonian is
\begin{equation}\begin{aligned}
\hat{H}
&=\frac{1}{2}
\begin{pmatrix}
\phi & \pi
\end{pmatrix}
\begin{bmatrix}
\text{Re}(A-B^*) & -\text{Im}(A)+\text{Im}(B)\\
\text{Im}(A)+\text{Im}(B)& \text{Re}(A+B) \\
\end{bmatrix}
\begin{pmatrix}
\phi\\
\pi
\end{pmatrix} \\
&=\frac{1}{2}
\begin{pmatrix}
\phi & \pi
\end{pmatrix}
\mathcal{M}
\begin{pmatrix}
\phi\\
\pi
\end{pmatrix}.
\end{aligned}\end{equation}
It is not hard to check that $\mathcal{M}$ is real and symmetric.

@@

We will not elaborate on Williamson's theorem and its proof (see proofs in Ref.~\onlinecite{xiao_theory_2009,pirandola_correlation_2009,gosson_symplectic_2006} and also a recent application in the entanglement entropy context\cite{coser_contour_2017}). Instead we will show in App.~\ref{app_sub:harmonic_chain} that for the problem of the harmonic chain we are interested in, the diagonalization can be easily done without using the general recipe in the Williamson theorem.

@@

Suppose we have obtained the required matrix $S$, the ground state will be the vacuum of the annihilation operators defined in Eq.~\eqref{eq:def_a} and in the $b$ basis it satisfies
\begin{equation}\begin{aligned}
\label{eq:a_vacuum_condition}
(b_iu_{ij}+b_i^\dagger v_{ij})|0\rangle_{{\bf a}}=0.
\end{aligned}\end{equation}
If the matrix $u$ is invertible, then we can introduce a matrix $T=vu^{-1}$ to rewrite Eq.~\eqref{eq:a_vacuum_condition} as
\begin{equation}\begin{aligned}
(b_i+b_j^\dagger T_{ji})|0\rangle_{{\bf a}}=0. 
\end{aligned}\end{equation}
The constraint Eq.~\eqref{eq:constraint_2} on the blocks of $u$ and $v$ (followed by the symplectic constraint of $S$) implies that $T$ is a symmetric matrix. With the observation of 
\begin{equation}\begin{aligned}
\exp\left\{-\frac{1}{2}b_j^\dagger T_{jk}b_k^\dagger\right\}b_i\exp\left\{\frac{1}{2}b_j^\dagger T_{jk}b_k^\dagger\right\}=b_i+T_{ij}b^\dagger_j,
\end{aligned}\end{equation}
we solve the groundstate
\begin{equation}\begin{aligned}
\label{eq:boson_BCS_gnd}
|0\rangle_{{\bf a}}&=\text{det}^{\frac{1}{4}}(1-T^\dagger T)\exp\left\{-\frac{1}{2}b_j^\dagger T_{jk}b_k^\dagger\right\}|0\rangle_{\bf b},\\
\end{aligned}\end{equation}
where the normalization is given by the McMahon master theorem Eq.~\eqref{eq:bosonic_McMahon}. Apply constraint in Eq.~\eqref{eq:constraint_1}, it simplifies to the top left corner of the symplectic matrix
\begin{equation}
\text{det}^{\frac{1}{4}}(1-T^\dagger T) =|\text{det}(u)|^{-\frac{1}{2}}.
\end{equation}

@@

We start with a quadratic bosonic Hamiltonian $\hat{H}_0$ in the ${\bf b}$ basis, as in Eq.~\eqref{eq:quadratic_boson_H}. From the discussion in App.~\ref{app_sub:boson_BdG}, we are able to diagonalize it in the ${\bf a}$ basis for positive definite $M$. At $t=0$, the Hamiltonian becomes $\hat{H}_1$, which is still written in the ${\bf b}$ basis, but is diagonalized in a new basis ${\bf c}$. The corresponding Bogoliubov transformations read
\begin{equation}\begin{aligned}
\label{eq:two_BdG}
({\bf b} , {\bf b}^\dagger)S_0=({\bf a} , {\bf a}^\dagger),\quad
({\bf b} , {\bf b}^\dagger)S_1=({\bf c} , {\bf c}^\dagger),
\end{aligned}\end{equation}
and so
\begin{equation}\begin{aligned}
\label{eq:S0invS}
({\bf a} , {\bf a}^\dagger)=({\bf c} , {\bf c}^\dagger)\left(S_0^{-1}S_1\right)^{-1}.
\end{aligned}\end{equation}
One realizes that Eq.~\eqref{eq:S0invS} is another Bogoliubov transformation and so the corresponding matrix has the block structure
\begin{equation}
S_1^{-1}S_0=\begin{bmatrix}
u_1 & v_1^*\\
v_1 & u_1^*
\end{bmatrix}.
\end{equation}
Thus $|0\rangle_{\bf c}$ is related to the $|0\rangle_{{\bf a}}$ in the same way as in Eq.~\eqref{eq:boson_BCS_gnd}. Their overlap is therefore given by the normalization factor
\begin{equation}\begin{aligned}
|{}_{\bf a}\langle0|0\rangle_{\bf c}|^2=\frac{\Big|{}_{\bf a}\langle 0 | \exp( -\frac{1}{2} a_j^{\dagger} T^{jk} a_k^{\dagger} )|0   \rangle_{\bf a} \Big|^2}{|\text{det}(u_1)|} = \frac{1}{|\text{det}(u_1)|}.
\end{aligned}\end{equation}

@@

In this subsection, we explicitly construct the matrix $\mathcal{S}$ for the case of the harmonic chain introduced in Sec.~\ref{sec_sub:free_boson_lattice}. In the basis defined in Eq.~\eqref{eq:real_basis}, the Hamiltonian for 1D harmonic chain is
\begin{equation}\begin{aligned}
\hat{H}
=\frac{1}{2}
\begin{pmatrix}
\phi & \pi
\end{pmatrix}
\mathcal{M}
\begin{pmatrix}
\phi\\
\pi
\end{pmatrix}
=\frac{1}{2}
\begin{pmatrix}
\phi & \pi
\end{pmatrix}
\begin{bmatrix}
\mathcal{V} \\
& {\mathbb{ I}}
\end{bmatrix}
\begin{pmatrix}
\phi\\
\pi
\end{pmatrix},
\end{aligned}\end{equation}
where $\mathcal{V}$ is real symmetric matrix that can be diagonalized as $\mathcal{V}=\mathcal{O}D^2\mathcal{O}^T$. The matrix $\mathcal{V}$ depends on the boundary condition, but positive definiteness is the only requirement here.

The matrix $\mathcal{S}$ that diagonalizes $\mathcal{M}$
\begin{equation}
\begin{aligned}
\mathcal{M}=\mathcal{S}
\begin{bmatrix}
D \\ 
& D
\end{bmatrix}
\mathcal{S}^T
\end{aligned}
\end{equation}
is given by the following real symplectic matrix
\begin{equation}\begin{aligned}
\mathcal{S}\equiv
\begin{bmatrix}
\mathcal{O}D^{1/2} \\
& \mathcal{O}D^{-1/2}
\end{bmatrix}.
\end{aligned}\end{equation}
